---
title: "Appendices"
author: "Tej Seth and Projeet Mukherjee"
output: html_document
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
```


```{r packages, include = FALSE}
library(tidyverse)
library(ggthemes)
library(car)
library(vip)
library(leaps)
library(HH)
library(qqplotr)
options(scipen = 9999)

theme_reach <- function() {
  theme_fivethirtyeight() +
    theme(
      legend.position = "none",
      plot.title = element_text(size = 22, hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(size = 18, hjust = 0.5),
      axis.title.x = element_text(size=18),
      axis.title.y = element_text(size=18),
      axis.text = element_text(size = 14),
      strip.text = element_text(size = 16, face = "bold"),
      legend.text = element_text(size = 14),
    )
}

```


```{r readData, include = FALSE}
fifa_22 <- read_csv(url("https://raw.githubusercontent.com/tejseth/stats-401-project/master/players_fifa22.csv"))

fifa_small <- fifa_22 %>%
  dplyr::select(FullName, Age, Height, Weight, PhotoUrl, Overall, BestPosition, ValueEUR,
         ShootingTotal, PassingTotal, DefendingTotal, PhysicalityTotal)

fifa_smallest <- sample_n(fifa_small, 4815)

fifa_smallest <- fifa_smallest %>% filter(ValueEUR > 0)
```

***

#### Figure 1: Histogram of Overall

```{r figure1, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = Overall)) +
  geom_histogram(fill = "#012e67", alpha = 0.9, bins = 40) +
  scale_color_identity() +
  theme_reach() +
  labs(x = "Overall Rating",
       y = "Count",
       title = "Histogram of Overall Rating in FIFA 22") +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```

***

#### Figure 2: Histogram of Shooting Rating

```{r figure2, echo=FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = ShootingTotal)) +
  geom_histogram(fill = "#012e67", alpha = 0.9, bins = 40) +
  scale_color_identity() +
  theme_reach() +
  labs(x = "Shooting Rating",
       y = "Count",
       title = "Histogram of Shooting Rating in FIFA 22") +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```

***

#### Figure 3: Histogram of Passing Rating

```{r figure3, echo=FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = PassingTotal)) +
  geom_histogram(fill = "#012e67", alpha = 0.9, bins = 40) +
  scale_color_identity() +
  theme_reach() +
  labs(x = "Passing Rating",
       y = "Count",
       title = "Histogram of Passing Rating in FIFA 22") +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```

***

#### Figure 4: Histogram of Defending Rating

```{r figure4, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = DefendingTotal)) +
  geom_histogram(fill = "#012e67", alpha = 0.9, bins = 40) +
  scale_color_identity() +
  theme_reach() +
  labs(x = "Defending Rating",
       y = "Count",
       title = "Histogram of Defending Rating in FIFA 22") +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```

***

#### Figure 5: Histogram of Physicality Total

```{r figure 5, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = PhysicalityTotal)) +
  geom_histogram(fill = "#012e67", alpha = 0.9, bins = 60) +
  scale_color_identity() +
  theme_reach() +
  labs(x = "Physicality Rating",
       y = "Count",
       title = "Histogram of Physicality Rating in FIFA 22") +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```

***

#### Figure 6: Histogram of Age

```{r figure6, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = Age)) +
  geom_histogram(fill = "#012e67", alpha = 0.9, bins = 40) +
  scale_color_identity() +
  theme_reach() +
  labs(x = "Age",
       y = "Count",
       title = "Histogram of Age in FIFA 22") +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```
***

#### Figure 7: Histogram of Height

```{r figure7, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = Height)) +
  geom_histogram(fill = "#012e67", alpha = 0.9, bins = 40) +
  scale_color_identity() +
  theme_reach() +
  labs(x = "Height",
       y = "Count",
       title = "Histogram of Height in FIFA 22") +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```

***

#### Figure 8: Histogram of Contract Value

```{r figure8, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = ValueEUR)) +
  geom_histogram(fill = "#012e67", alpha = 0.9, bins = 40) +
  scale_color_identity() +
  theme_reach() +
  labs(x = "Contract Value (Euro's)",
       y = "Count",
       title = "Histogram of Contract Value in FIFA 22") +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_y_continuous(breaks = scales::pretty_breaks())
```

*** 

#### Figure 9: Scatter Plot of Contract Value and Overall

```{r figure9, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = log(ValueEUR), y = Overall)) +
  geom_smooth(method = "lm", se = FALSE, color = "black", size = 1.2) +
  geom_point(fill = "darkorange", shape = 21, color = "black", alpha = 0.8, size = 3.5) +
  theme_reach() +
  labs(x = "Log-Transformed Contract Value (Euros)",
       y = "Overall (0-100)",
       title = "Scatterplot of Logged Contract Value and Overall Rating")
```
*** 

#### Figure 10: Scatter Plot of Age and Overall

```{r figure10, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = Age, y = Overall)) +
  geom_smooth(method = "lm", se = FALSE, color = "black", size = 1.2) +
  geom_point(fill = "darkorange", shape = 21, color = "black", alpha = 0.8, size = 3.5) +
  theme_reach() +
  labs(x = "Age (Years)",
       y = "Overall (0-100)",
       title = "Scatterplot of Age and Overall Rating")
```

*** 

#### Figure 11: Scatter Plot of Height and Overall

```{r figure11, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = Height, y = Overall)) +
  geom_smooth(method = "lm", se = FALSE, color = "black", size = 1.2) +
  geom_point(fill = "darkorange", shape = 21, color = "black", alpha = 0.8, size = 3.5) +
  theme_reach() +
  labs(x = "Height (Centimeters)",
       y = "Overall (0-100)",
       title = "Scatterplot of Height and Overall Rating")
```

*** 

#### Figure 12: Scatter Plot of Shooting Rating and Overall

```{r figure12, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = ShootingTotal, y = Overall)) +
  geom_smooth(method = "lm", se = FALSE, color = "black", size = 1.2) +
  geom_point(fill = "darkorange", shape = 21, color = "black", alpha = 0.8, size = 3.5) +
  theme_reach() +
  labs(x = "Shooting Rating (0-100)",
       y = "Overall (0-100)",
       title = "Scatterplot of Shooting Rating and Overall Rating")
```

*** 

#### Figure 13: Scatter Plot of Defending Rating and Overall

```{r figure13, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = DefendingTotal, y = Overall)) +
  geom_smooth(method = "lm", se = FALSE, color = "black", size = 1.2) +
  geom_point(fill = "darkorange", shape = 21, color = "black", alpha = 0.8, size = 3.5) +
  theme_reach() +
  labs(x = "Defending Rating (0-100)",
       y = "Overall (0-100)",
       title = "Scatterplot of Defending Rating and Overall Rating")
```

*** 

#### Figure 14: Scatter Plot of Passing Rating and Overall

```{r figure14, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = PassingTotal, y = Overall)) +
  geom_smooth(method = "lm", se = FALSE, color = "black", size = 1.2) +
  geom_point(fill = "darkorange", shape = 21, color = "black", alpha = 0.8, size = 3.5) +
  theme_reach() +
  labs(x = "Passing Rating (0-100)",
       y = "Overall (0-100)",
       title = "Scatterplot of Passing Rating and Overall Rating")
```

*** 

#### Figure 15: Boxplot of Overall Rating by Position

```{r figure15, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  group_by(BestPosition) %>%
  mutate(med = median(Overall)) %>%
  ungroup() %>% 
  ggplot(aes(x = fct_reorder(BestPosition, med), y = Overall)) +
  geom_boxplot(aes(fill = fct_reorder(BestPosition, med))) +
  scale_fill_viridis_d() +
  theme_reach() +
  labs(x = "Position",
       y = "Overall Rating",
       title = "Overall Rating by Position in Fifa 22",
       fill = "Position") +
  theme(axis.text.x = element_text(angle = -30))
```

*** 

#### Figure 16: Interaction Terms

```{r figure16, echo = FALSE, fig.height = 10, fig.width = 16}
fifa_smallest %>%
  ggplot(aes(x = log(ValueEUR), y = Overall, color = BestPosition)) +
  geom_point(aes(shape = BestPosition, color = BestPosition), size = 4, alpha = 0.2) +
  geom_smooth(se = FALSE, size = 2) +
  theme_reach() +
  theme(legend.position = "bottom") +
  labs(x = "Log-Transformed Contract Value (Euros)",
       y = "Overall (0-100)",
       title = "Scatterplot of Logged Contract Value and Overall Rating by Position")
```

*** 

#### Figure 17: Building the Model Subsets

```{r figure17, echo = FALSE}
fifa_full <- regsubsets(Overall ~ Age + Height + ShootingTotal + I(ShootingTotal^2) + 
                 DefendingTotal + I(DefendingTotal^2)  + PassingTotal + 
                PhysicalityTotal + log(ValueEUR) * BestPosition, data = fifa_smallest)

summaryHH(fifa_full)
```

*** 

#### Figure 18: VIF Scores for the Model

```{r figure18, echo = FALSE}
lm_fifa <- lm(Overall ~ Age + Height + ShootingTotal + I(ShootingTotal^2) + 
                 DefendingTotal + I(DefendingTotal^2)  + PassingTotal + 
                PhysicalityTotal + log(ValueEUR) * BestPosition, data = fifa_smallest)
                 
#summary(lm_fifa)

lm_fifa_vif <- lm(Overall ~ Age + Height +
                 DefendingTotal  + PassingTotal + PhysicalityTotal + 
                   log(ValueEUR) + BestPosition, data = fifa_smallest)

vif(lm_fifa_vif)

```

*** 

#### Figure 19: Final Model

```{r figure19, echo = FALSE}
fifa_full <- lm(Overall ~ Age + Height + DefendingTotal + 
                        I(DefendingTotal^2)  + PassingTotal + PhysicalityTotal + 
                          log(ValueEUR) * BestPosition, data = fifa_smallest)

summary(fifa_full)

```

*** 

#### Figure 20: Residuals vs. Fitted Plot

```{r figure20, echo = FALSE, fig.width = 16, fig.height = 10}
plot(fifa_full$fitted.values, fifa_full$residuals, 
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Fitted Value and Residuals Plot")

```

#### Figure 21: QQ Plot

```{r figure21, echo = FALSE, fig.width = 16, fig.height = 10}
qqnorm(fifa_full$residuals, main = "QQ Plot of the Residuals")
qqline(fifa_full$residuals)

```